A Continual Planner that Switches Between Serial and Decision-Theoretic Planning

### 2010-11-16 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Tiny changes
 - Charles

We developed a first-order declarative language, called DTPDDL, for
describing domains of planning problems that correspond to POMDPs.  To
accompany that language we have developed an information-state
\laostar\ procedure for solving problems expressed in DTPDDL. We have
an automated determinisation procedure that generates MAPL
descriptions given DTPDDL input. To accompany that, we have extended
MAPSIM to simulate DTPDDL instances, and modified \fastdownward\ so
that it can find useful serial executions given DTPDDL and MAPL models
of the problem at hand.

@@

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. The underlying world
dynamics are modelled in terms of a finite set of stochastic actions
$\stochActions$, deterministic outcomes $\detActions$, and
state-characterising propositions $\propositions$.  A problem state
$\state$ is a set of propositions $\state \subseteq \props$, and we
denote \states\ the set of reachable states. The perceptual model is
given in terms of a finite set of stochastic senses $\stochSenses$,
deterministic sensing outcomes $\detSenses$, and perceptual
propositions $\percepts$, called percepts. Here, an observation
$\observ$ is a set of percepts $\observ \subseteq \percepts$, and we
denote \observations\ the set of reachable observations.

Every stochastic action $\stochAction \in \stochActions$ has a
precondition $\poss(\stochAction)$, which is a set of propositions. An
action is applicable at a state $\state$ when $\poss(\stochAction)
\subseteq \state$. We denote $\stochActions(\state)$ the set of
actions applicable at state $\state$.  When $\stochAction \in
\stochActions(\state)$ is applied, nature chooses one elements amongst
a small set of deterministic outcomes $\detActions(\stochAction)
\equiv \{\detAction_1, \ldots, \detAction_k \}$. We denote
$\mu_{\stochAction}(\detAction_i)$ the probability that nature takes
outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. That outcome so chosen has an
effect on the underlying state given in terms of two lists of
propositions. The add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
$\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
..,\prop_n]$ and $\delete(\detAction) := [\prop_1, ..,\prop_m]$ is
chosen by nature when \stochAction\ is applied at state $\state$, then
the resultant state is $ ( \state \cup \add(\detAction) ) \backslash
\delete(\detAction)$ -- i.e., propositions from $\add(\detAction)$ are
added to $\state$, and those from $\delete(\detAction)$ are removed
from $\state$.

We are concerned with problems that feature partial
observability. Here the underlying state of the world cannot be
observed directly, rather, senses $\stochSense \in \stochSenses$
effect an observation $\observ \in \observations$ that informs what
should be believed about the world. In the POMDP model, if $\action$
is applied effecting a transition to a successor state $\state'$, then
an observation occurs according to the active senses
$\stochSenses(\stochAction, \state') \subseteq \stochSenses$. A
precept is active, written $\stochSense \in
\stochSenses(\stochAction, \state')$, if the senses
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition
$\poss_\states(\stochSense) \subseteq \propositions$ is satisfied by
the state $\state'$; In the usual sense that
$\poss_\states(\stochSense) \subseteq \state'$.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from chosen elements of active senses.

@@

Whereas until now we have considered a POMDP model factored in terms
of propositions and percepts, in order to discuss utilities and
policies it is convenient to consider the underlying decision process
in a flat format. This underlying decision process is given by the
tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
belief-state, \states\ is the finite set of states, \actions\ is the
finite set of actions, and \observations\ is the finite set of
observations.  Where $\state, \state' \in \states$, $a \in \actions$,
from $\mu$ we have a state transition function $\transProb(\state,
\action, \state')$ giving the probability of a transition from state
$s$ to $s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
Function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function.  Consequently there is a positive constant $c$
so that for all $\state \in \states$ and $\action \in \actions$,
$|\reward(\state, \action)| < c$.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.

>>>>

We developed a first-order declarative language, called DTPDDL, for
describing domains of planning problems that correspond to POMDPs.  To
accompany that language we have implemented an information-state
\laostar\ procedure for solving problems expressed in DTPDDL. We have
an automated determinisation procedure that generates MAPL
descriptions given DTPDDL input. To accompany that, we have extended
MAPSIM to simulate DTPDDL instances, and modified \fastdownward\ so
that it can find useful serial executions given DTPDDL and MAPL models
of the problem at hand.

@@

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. The underlying world
dynamics are modelled in terms of a finite set of stochastic actions
$\stochActions$, deterministic outcomes $\detActions$, and
state-characterising propositions $\propositions$.

A problem state $\state$ is a set of propositions $\state \subseteq
\props$. Every stochastic action $\stochAction \in \stochActions$ has
a precondition $\poss(\stochAction)$, also a set of propositions. An
action is applicable at a state $\state$ when $\poss(\stochAction)
\subseteq \state$. We denote $\stochActions(\state)$ the set of
actions applicable at state $\state$.  When $\stochAction \in
\stochActions(\state)$ is applied, nature chooses one element amongst
a small set of deterministic outcomes $\detActions(\stochAction)
\equiv \{\detAction_1, \ldots, \detAction_k \}$. We denote
$\mu_{\stochAction}(\detAction_i)$ the probability that nature takes
outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. That outcome so chosen has an
effect on the underlying state given in terms of two lists of
propositions. The add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
  $\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
  ..,\prop_n]$ and $\delete(\detAction) := [\prop_1, ..,\prop_m]$ is
chosen by nature when \stochAction\ is applied at state $\state$, then
the resultant state is $ ( \state \cup \add(\detAction) ) \backslash
\delete(\detAction)$ -- i.e., propositions from $\add(\detAction)$ are
added to $\state$, and those from $\delete(\detAction)$ are removed
from $\state$.

We are concerned with problems that feature partial
observability. These have a perceptual model given in terms of a
finite set of stochastic senses $\stochSenses$, deterministic sensing
outcomes $\detSenses$, and perceptual propositions $\percepts$, called
percepts. Here, an observation $\observ$ is a set of percepts $\observ
\subseteq \percepts$, and we denote \observations\ the set of
reachable observations. The underlying state of the world cannot be
observed directly, rather, senses $\stochSense \in \stochSenses$
effect an observation $\observ \in \observations$ that informs what
should be believed about the world. In the POMDP model, if $\action$
is applied effecting a transition to a successor state $\state'$, then
an observation occurs according to the active senses
$\stochSenses(\stochAction, \state') \subseteq \stochSenses$. A
precept is active, written $\stochSense \in \stochSenses(\stochAction,
\state')$, if the senses action-precondition,
$\poss_\stochActions(\stochSense)$, is equal to $\stochAction$, and
the state-precondition $\poss_\states(\stochSense) \subseteq
\propositions$ is satisfied by the state $\state'$; In the usual sense
that $\poss_\states(\stochSense) \subseteq \state'$.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from chosen elements of active senses.

@@

Whereas until now we have considered a POMDP model factored in terms
of propositions and percepts, in order to discuss utilities and
policies it is convenient to consider the underlying decision process
in a flat format. This underlying decision process is given by the
tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations.  Where $\state, \state' \in
\states$, $a \in \actions$, from $\mu$ we have a state transition
function $\transProb(\state, \action, \state')$ giving the probability
of a transition from state $s$ to $s'$ if $a$ is applied. For any
$\state$ and $\action$ we have $\sum_{\state' \in \states}
\transProb(\state, \action, \state') = 1$.
%%
Function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function.  Consequently there is a positive constant $c$
so that for all $\state \in \states$ and $\action \in \actions$,
$|\reward(\state, \action)| < c$.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.
